---
title: "Washington"
author: "Tianyu Zhang"
date: "`r Sys.Date()`"
output: 
  html_document:
    keep_md: yes
    toc: yes
    theme: flatly
    toc_float: yes
    code_folding: hide
    number_sections: no
---

<style>
.kable thead tr th, .table thead tr th {
  text-align: left !important;}
table.kable, table.table {
  width: 100% !important;}
  body {
  line-height: 1.6;
  font-size: 16px
}
</style>

# Load Data

```{r}
if(!require(pacman)){install.packages("pacman"); library(pacman)}
p_load(tidyverse, sf, units)

setwd("C:/Users/ztyuu/Documents/Spring2024/Research/Ghost Forest")

Washington.sf <- st_read("Washington/Parcels/Parcels.shp")

Washington.sf <- Washington.sf %>% 
  rename(PIN = Parcels_NC)
```

# Processing Parcel Records

## Eliminating outliers

Parcels satisfying the following qualities are removed from the Washington County dataset:

- PIN is NA: These parcels are not associated with any values. **Note: The only NA parcel with land value has Parcels_OB=3077. It is removed for now to keep the logic consistent**

```{r}
Washington.sf <- Washington.sf %>% 
  filter(!is.na(PIN))
```

# Check for real *geographical* duplicates based on geometry

```{r check PIN duplicates}
W.geo.asdf <- Washington.sf %>%
  as_tibble()%>%
  group_by(geometry) %>% 
  tally() %>% 
  filter(n>1)
```

## Extracting real *geographical* duplicates based on geometry

```{r}
W.realdups <- W.geo.asdf %>% 
  select(-n) %>% 
  inner_join(Washington.sf, by = "geometry") %>% 
  distinct(geometry, .keep_all = TRUE) %>% 
  mutate(Duped_Value = "0")
```

# Check for *value* duplicates based on PIN and ownership

Washington county dataset is unique as in value records are duplicated in two ways - by ownership and by PIN. The following steps treats them differently.

- Value duplicates by ownership occur across three cemeteries: ZIONS CHAPEL CHURCH OF CHRIST, HILLSIDE MEMORIAL GARDENS, WINDLEY CEMETERY.

- Value duplicates by PIN occur across parcels with the same PIN.

## Cleaning *value* duplicates (fake geographical duplicates) by ownership - three cemeteries

### Cleaning *value* duplicates (fake geographical duplicates) - ZIONS CHAPEL CHURCH OF CHRIST

All parcels with Pin formatted 2 or 3 digits are cemetery units owned by Zion Chapel. WALK are corridors owned by Zion Chapel. These parcels' total assessed values are collectively 9,700 and land acres 0.69, suggesting they represent the same deed record. Since 0.69 acres is considerably large, it is most likely 9,700 is the total value associated with the entire cemetery. 

The ZIONS CHAPEL CHURCH OF CHRIST parcels are extracted and grouped together under a common PIN "ZionsChapel".

```{r}
W.zion <- Washington.sf %>%
  filter(Parcels__4 == "ZION CHURCH CEMETERY") %>% 
  mutate(PIN = "ZionsChapel",
         Duped_Value = "1")
```

### Cleaning fake *value* duplicates (fake geographical duplicates) - HILLSIDE MEMORIAL GARDENS

```{r}
W.hillside <- Washington.sf %>%
  filter(Parcels__4 == "HILLSIDE MEMORIAL GARDENS") %>% 
  mutate(PIN = "HillsideMemorial",
         Duped_Value = "1")
```

### Cleaning *value* duplicates (fake geographical duplicates) - WINDLEY CEMETERY

```{r}
W.windley <- Washington.sf %>%
  filter(Parcels__4 == "WINDLEY CEMETERY") %>% 
  mutate(PIN = "Windley",
         Duped_Value = "1")
```

## Cleaning *value* duplicates (fake geographical duplicates) by PIN

For other fake duplicates:

- Value of each record is the total of all parcels under the same PIN.

- Owner info is consistent for all parcels under the same PIN.

```{r}
W.fakedups <- Washington.sf %>%
  as_tibble()%>%
  group_by(PIN) %>% 
  tally() %>% 
  filter(n>1) %>% 
  select(-n) %>%
  inner_join(Washington.sf, by = 'PIN') %>% 
  anti_join(W.zion, by = 'geometry') %>%  #this step is manually repeated to ensure no dups are produced in data cleansing
  anti_join(W.hillside, by = 'geometry') %>% 
  anti_join(W.windley, by = 'geometry') %>% 
  anti_join(W.realdups, by = 'geometry') %>% 
  distinct(geometry, .keep_all = TRUE) %>% 
  mutate(Duped_Value = "0")
```

## Identifying unique parcels

```{r}
W.unique <- Washington.sf %>% 
  as_tibble() %>% 
  anti_join(W.zion, by = "geometry") %>% 
  anti_join(W.hillside, by = 'geometry') %>% 
  anti_join(W.windley, by = 'geometry') %>% 
  anti_join(W.realdups, by = 'geometry') %>% 
  anti_join(W.fakedups, by = 'geometry') %>% 
  mutate(Duped_Value = "0")
```


# Processing Master Shapefile

```{r}
WashingtonMaster <- rbind(W.unique, W.realdups, W.fakedups, W.hillside,W.windley, W.zion) %>% 
  st_as_sf()
```

```{r}
WashingtonMaster <- WashingtonMaster %>%
  mutate(Owner1 = paste(Parcels_OW, Owner)) %>% 
  #paste two columns of owner info to retain both individual and company details
  select(PIN, Owner1, Parcels__2, Parcels_CI, Parcels_ST, Parcels_ZI, 
         LandAssess, BldgAssess, DetachedAs, Duped_Value) %>% 
  rename(Land_Val = LandAssess, Struc_Val = BldgAssess, Other_Val = DetachedAs,
         Mail_St = Parcels__2, Mail_City = Parcels_CI, Mail_State = Parcels_ST,
         Mail_ZIP = Parcels_ZI) %>% 
  mutate(Owner2 = NA_character_,
         PIN = as.character(PIN),
         Land_Val = as.numeric(Land_Val),
         Struc_Val = as.numeric(Struc_Val),
         Other_Val = as.numeric(Other_Val),
         Total_Val = Land_Val + Struc_Val + Other_Val,
         Acres = st_area(.),
         Acres = set_units(x = Acres, value = "acres"),
         Total_Val = as.numeric(Total_Val),
         Mail_ZIP = as.character(Mail_ZIP),
         Parcel_County = "Washington",
         ID = row_number(),
         ID = paste0("W", sprintf("%05d", ID)))

#manually removed one last geometry duplicate

WashingtonMaster <- WashingtonMaster %>% 
  filter(ID != 'W11133')
```

# Export Final Shapefile

```{r}
if (file.exists("WashingtonMaster.shp")) {
  file.remove("WashingtonMaster.shp")
}
st_write(WashingtonMaster, "WashingtonMaster.shp", delete.dsn = TRUE)
```